Reducing iptables configuration complexity using chains

### 2014-05-04 iptables/paper.tex
improve conclusion
add brief introduction to iptables and chains section

\section{Brief introduction to iptables}
\subsection{ESTABLISHED state}
When using this option, you can filter for established connections. If you define it in both the INPUT and OUTPUT rules, you only have to define in the INPUT rules which NEW incoming requests should be allowed, and in the OUTPUT rules which NEW outgoing request are allowed. The established connections will be allowed and should not be redefined (making the configuration a lot more readable and maintainable). An example allowing only an SSH service without using the ESTABLISHED state would be :
\begin{verbatim}
  iptables -A INPUT -p tcp --dport ssh -j ACCEPT
  iptables -A INPUT -j REJECT
  iptables -A OUTPUT -p tcp --sport ssh -j ACCEPT
  iptables -A OUTPUT -j REJECT
\end{verbatim}

@@

\section{Conclusion}
\begin{itemize}
  \item When checking for the ESTABLISHED state, the number of rules can be reduced
  \item Using chains makes your rules better structured which makes them easier to read and better maintainable
  \item Chains can be reused for several rules
  \item Chains can be chained together
  \item Using chains makes checking rules faster, because it only jumps to a chain when a rule matches. Not all rules have to be checked.
\end{itemize}

>>>>

\section{Brief introduction to iptables}
  Iptables is a tool for creating the rulesets for netfilter, a packet filtering framework which was introduced in the linux 2.4 kernel.\\

  Netfilter is rule based. When an IP packet arrives, it is checked against a set of rules, fe. :
  \begin{verbatim}
    iptables -A INPUT -m tcp -p tcp \dd dport ssh -j ACCEPT
    iptables -A INPUT -m tcp -p tcp \dd dport http -j ACCEPT
    iptables -A INPUT -m tcp -p tcp \dd dport https -j ACCEPT
    iptables -A INPUT -j DROP
  \end{verbatim}

  Every rule consists of several parts :
  \begin{itemize}
    \item the chain the rule belongs to
    \item parameters filtering for a type of packet
    \item the target : action to be taken when a packet matches
  \end{itemize}

  \subsection{Chains}
    There are several predefined chains :
    \begin{itemize}
      \item INPUT
      \item OUTPUT
      \item FORWARD
      \item PREROUTING
      \item POSTROUTING
    \end{itemize}
    Defining custom chains is also possible.\\

    In this example, the rule is added to the INPUT chain :
    \begin{verbatim}
iptables -A INPUT -m tcp -p tcp --dport ssh -j ACCEPT
    \end{verbatim}

\subsection{ESTABLISHED state}
When using this option, you can filter for established connections. If you define it in both the INPUT and OUTPUT rules, you only have to define in the INPUT rules which NEW incoming requests should be allowed, and in the OUTPUT rules which NEW outgoing request are allowed. The established connections will be allowed and should not be redefined (making the configuration a lot more readable and maintainable). An example allowing only an SSH service without using the ESTABLISHED state would be :
\begin{verbatim}
  iptables -A INPUT -p tcp --dport ssh -j ACCEPT
  iptables -A INPUT -j REJECT
  iptables -A OUTPUT -p tcp --sport ssh -j ACCEPT
  iptables -A OUTPUT -j REJECT
\end{verbatim}

@@

\section{Conclusion}
\begin{itemize}
  \item When checking for the ESTABLISHED state, the number of rules can be reduced, because only the initiating packet needs to be checked
  \item Using chains makes your rules better structured which makes them easier to read and better maintainable
  \item Chains can be reused for several rules
  \item Chains can be chained together
  \item Using chains makes checking rules faster, because it only jumps to a chain when a rule matches. Not all rules have to be checked
\end{itemize}
